\batcave: Bayesian Analysis Tools for Context-Aware Variant Evaluation

### 2019-08-17 NAR_DRAFT.tex
trying to add supplement. not easy

@@

\begin{figure}[t]
  \begin{center}
  \includegraphics{figures/signature_figure.pdf}
  \end{center}
  \caption{Mutation profiles. 
  \textbf{(a)} The observed mutation profile of a breast cancer sample from \cite{Shi2018}.
  \textbf{(b)} The observed mutation profile of the AML 31 primary tumor from \cite{Griffith2015}.
  \textbf{(c)} A mutation profile used for simulating tumors, made up of equal proportions of COSMIC mutation signatures 1, 7, \& 11.
  \textbf{(d)} Equal proportions of signatures 1, 4, \& 5.
  \textbf{(e)} Equal proportions of signatures 1, 3, \& 5.
  KL divergence between the estimated profile ($\boldsymbol{\pi}$) and the simulated mutation profile for \textbf{(f)} Whole exomes and \textbf{(g)} whole genomes.
  The estimated mutation profile converges quickly to the target as the number of mutations evaluated rises.
  }
  \label{NAR-sigfig}
  \end{figure}

@@

@@

\section{RESULTS}
We designed \batcave as a post-call variant evaluation algorithm to be used with MuTect (Versions 1.1.7 or $>$2.0) \cite{Cibulskis2013}.
The algorithm is effecient, processing 22,000 variants per second on a typical iMac desktop computer, which corresponds to 100 seconds to process a 500X exome and 2000 seconds for a 100X whole genome.
MuTect is widely used, has state of the art sensitivity and specificity, and includes numerous heuristic filters and alignment adjustments that reduce the prevalence of sequencing errors in results \cite{Griffith2015}.
\batcave takes the output from MuTect, extracts the log-likelihood ratio for each potential variant site, and uses that ratio to separate the potential sites in to high-confidence and low-confidence groups.
The mutation profile is computed from the high confidence sites, and the posterior probability of mutation is recomputed for all sites as described in Materials and Methods.
\begin{figure*}
  \begin{center}
  \includegraphics{figures/fig_wes.pdf}
  \end{center}
  \caption{Model performance on 500X simulated whole exome.
  \textbf{(a)} Precision recall curves and \textbf{(c)} Receiver operating characteristic curves for 3 mutation signatures.
  Distribution of estimated true positive probabilities for true positive (top) and true negative (bottom) variants for \textbf{b)} the MuTect model and \textbf{d)} the \batcave model.
  A perfectly calibrated model would generate the diagonal line.}
\label{NAR-wes_fig}
\end{figure*}

@@

\begin{figure}
  \begin{center}
  \includegraphics{figures/ppv_wes.pdf}
  \end{center}
  \caption{Model performance on 500X simulated whole exome. 
  Positive predictive values vs. Recall for three different probability cutoffs. 
  Note the restricted axes to improve readability of points.}
\label{NAR-ppv_fig}
\end{figure}

@@

@@

\section{Supplementary Figures}

% \begin{figure*}[t]
%   \begin{center}
%   \includegraphics{figures/aml_plot.pdf}
%   \end{center}
%   \caption{Model performance on 300X Whole leukemia genome.
%   \textbf{(a)} Precision/recall curves and \textbf{(c)} Reciever operating characteristic curves.
%   Distribution of estimated true positive probabilities for true positive (top) and true negative (bottom) variants for \textbf{b)} the MuTect model and \textbf{d)} the Calibrated model.
%   A perfectly calibrated model would generate the diagonal line.}
%   \label{NAR-aml}
% \end{figure*}

\begin{figure*}
  \begin{center}
  \includegraphics{figures/fig_wgs.pdf}
  \end{center}
  \caption{Model performance on 100X simulated whole genome.
  \textbf{(a)} Precision recall curves and \textbf{(c)} Reciever operating characteristic curves for 3 mutation signatures.
  Distribution of estimated true positive probabilities for true positive (top) and true negative (bottom) variants for \textbf{b)} the MuTect model and \textbf{d)} the \batcave model.
  A perfectly calibrated model would generate the diagonal line.\bkmcomment{change the axis on b and d to be the same as the steyerberg paper. much clearer}}
\label{NAR-wgs_fig}
\end{figure*}

\begin{table*}[b]
  \tableparts{%
  \caption{This is a table with $\mu$, auroc, auprc, ICI, and maybe fraction called?}
  \label{table:01}%
  }{%
  \begin{tabular*}{\textwidth}{@{}lllllllll@{}}
  \toprule
  Depth & Mutation profile & $\mu$ & AUROC & & AUPRC & & ICI & 
  \\
  & & (estimated) & MuTect & \batcave & MuTect & \batcave & MuTect & \batcave
  \\
  \colrule
  100X whole genome & Concentrated & 3.6e-7 & .987 & .993 & .972 & .975 & -- & --
  \\
  100X whole genome & Intermediate & 3.2e-7 & .987 & .989 & .972 & .973 & -- & --
  \\
  100X whole genome & Diffuse & 3.2e-7 & .988 & .989 & .971 & .973 & -- & --
  \\
  500X whole exome & Concentrated & 3.6e-7 & .848 & .929 & .674 & .758 & .138 & .109
  \\
  500X whole exome & Intermediate & 3.6e-7 & .847 & .881 & .677 & .706 & .108 & .112 
  \\
  500X whole exome & Diffuse & 3.6e-7 & .850 & .873 & .676 & .698 & .105 & .116
  \\
  \citet{Williams2016} & Actual & 3.6e-8 & -- & -- & .995 & .996 & -- & --
  \\
  \citet{Shi2018} & Actual & 3.6e-8 & -- & -- & .972 & .972 & -- & --
  \\
  \botrule
  \end{tabular*}%
  }
  {$\mu$ = per-base mutation rate, AUROC/AUPRC = area under ROC/PRC curve, ICI = Integrated calibration index.}
  \end{table*}

\begin{figure}
  \begin{center}
  \includegraphics{figures/ppv_wgs.pdf}
  \end{center}
  \caption{Model performance on 500X simulated whole exome. 
  Positive predictive values vs. Recall for three different probability cutoffs. 
  Note the restricted axes to improve readability of points.}
\label{NAR-ppv_wgs_fig}
\end{figure}

>>>>

\begin{figure}
  \begin{center}
  \includegraphics{figures/signature_figure.pdf}
  \end{center}
  \caption{Mutation profiles. 
  \textbf{(a)} The observed mutation profile of a breast cancer sample from \cite{Shi2018}.
  \textbf{(b)} The observed mutation profile of the AML 31 primary tumor from \cite{Griffith2015}.
  \textbf{(c)} A mutation profile used for simulating tumors, made up of equal proportions of COSMIC mutation signatures 1, 7, \& 11.
  \textbf{(d)} Equal proportions of signatures 1, 4, \& 5.
  \textbf{(e)} Equal proportions of signatures 1, 3, \& 5.
  KL divergence between the estimated profile ($\boldsymbol{\pi}$) and the simulated mutation profile for \textbf{(f)} Whole exomes and \textbf{(g)} whole genomes.
  The estimated mutation profile converges quickly to the target as the number of mutations evaluated rises.
  }
  \label{NAR-sigfig}
  \end{figure}

@@

@@

add to change

@@

  \begin{figure*}
    \begin{center}
    \includegraphics{figures/fig_wes.pdf}
    \end{center}
    \caption{Model performance on 500X simulated whole exome.
    \textbf{(a)} Precision recall curves and \textbf{(c)} Receiver operating characteristic curves for 3 mutation signatures.
    Distribution of estimated true positive probabilities for true positive (top) and true negative (bottom) variants for \textbf{b)} the MuTect model and \textbf{d)} the \batcave model.
    A perfectly calibrated model would generate the diagonal line.}
  \label{NAR-wes_fig}
  \end{figure*}

\section{RESULTS}
We designed \batcave as a post-call variant evaluation algorithm to be used with MuTect (Versions 1.1.7 or $>$2.0) \cite{Cibulskis2013}.
The algorithm is effecient, processing 22,000 variants per second on a typical iMac desktop computer, which corresponds to 100 seconds to process a 500X exome and 2000 seconds for a 100X whole genome.
MuTect is widely used, has state of the art sensitivity and specificity, and includes numerous heuristic filters and alignment adjustments that reduce the prevalence of sequencing errors in results \cite{Griffith2015}.
\batcave takes the output from MuTect, extracts the log-likelihood ratio for each potential variant site, and uses that ratio to separate the potential sites in to high-confidence and low-confidence groups.
The mutation profile is computed from the high confidence sites, and the posterior probability of mutation is recomputed for all sites as described in Materials and Methods.

@@

@@

\begin{figure}
  \begin{center}
  \includegraphics{figures/ppv_wes.pdf}
  \end{center}
  \caption{Model performance on 500X simulated whole exome. 
  Positive predictive values vs. Recall for three different probability cutoffs. 
  Note the restricted axes to improve readability of points.\bkmcomment{how do called allele frequency distributions differ?}}
\label{NAR-ppv_fig}
\end{figure}

@@

\include{supplement}
